FastSRM: A fast, memory efficient and identifiable implementation of the shared response model

### 2021-12-15 paper/main.tex
small pass

In this work, we implement an identifiable version of SRM and show on real data
that it improves the stability of the recovered shared response.
%
We then introduce the FastSRM that relies on a dimension reduction step that we
prove to yield the same solution as the original algorithm.
% be optimal in the sense that working with reduced data does not induce any change in the
% algorithm trajectory.
%
We show experimentally using synthetic and real fMRI data
that FastSRM is considerably faster and more memory efficient
than current implementations.

@@

\section{Introduction}
%\bt{We need to start with a more application-oriented intro}
%
Complex stimulations such as movie-watching, story or music listening are more
and more popular in the neuro-scientific community. Indeed such naturalistic
paradigms are unconstrained from behavioral manipulations and thus, more
ecological with respect to real-world conditions.
%
However and in contrast to more traditional designs, these so-called
naturalistic stimuli do not come with a design matrix that quantifies the
features of interest in the stimuli.
%
Without a design matrix, models such as the general linear
model~\cite{poline2012general} cannot be used. Although some works have used
natural annotations~\cite{huth2012continuous}, see also \url{https://neuroscout.org} or deep neural
networks~\cite{gucclu2017increasingly}~\cite{richard2018optimizing} to create an
estimate of the design matrix of naturalistic
stimuli it is by no means a simple task.

Another approach is to learn jointly the design matrix and the spatial maps in
an unsupervised way.
%
In the case where the dataset contains only one subject
independent component analysis~\cite{jutten1991blind} (ICA) is the method of choice.
%
ICA assumes that components are independent, a rather realistic assumption which
ensures identifiability up to permutation and scaling of the model and can be fitted efficiently.
%

But when the dataset contains multiple subjects, a natural assumption is to
assume that the design matrix is shared across subjects.
%
Many different methods
can produce shared components from different subjects. Some assume independent
components~\cite{richard2021model}~\cite{richard2020modeling}~\cite{varoquaux2009canica}~\cite{calhoun2001method}.
while others are only based on second order information such as Multiset
Canonical Correlation Analysis~\cite{via2011joint} or the Shared Response Model (SRM)~\cite{chen2015reduced}.
%

This work focuses on the SRM. Because of its
simplicity and its built-in dimension reduction, this model is widely
used~\cite{baldassano2017discovering}~\cite{cohen2017computational}~\cite{baldassano2018representation}~\cite{jolly2020custom}~\cite{lee2021anticipation}
including as a preprocessing step for source
identification~\cite{richard2021model}~\cite{richard2020modeling},
classification~\cite{turek2018capturing}~\cite{chen2017shared}\cite{zhang2016searchlight} or as a basis for transfer learning~\cite{zhang2018transfer}.

@@

In this work, we introduce FastSRM, an identifiable model that uses an optimal
spatial decomposition to speed up the computations with provably no loss of
performance.
%
FastSRM gains identifiability by imposing that the covariance of the shared
components is diagonal (see ~\cite{richard2020modeling} Appendix D for a proof).
%
The gains in speed and memory usage are based on the observation that a low-dimensional, yet distance-preserving
representation of the images yields the same result as the full data.
%
Such a representation can be interpreted as an initial spatial  decomposition of the data.
% that reduces the data in a such a way that the obtained
% shared components (and other parameters) are the same as if they were computed
% on full data.
%
We show that FastSRM brings several practical benefits: on real data, it produces more stable
estimates of the shared components than its non-identifiable counterpart and 
 it is much faster and more memory
efficient than other implementations that do not make use of an optimal spatial decomposition.
% on synthetic as well as on real data,

@@

\subsection{Deterministic shared response model}
\label{sec:deterministicsrm}
Let us consider $n$ observations of $\xb_i$ and $\sbb$ that we stack into
matrices $X_i \in \RR^{v, n}$ and $S \in \RR^{p, n}$.

The deterministic shared response model considers both the mixing matrices $A_i$ and
the $n$ observations of the shared response $S$ as parameters to be
estimated.
%
The noise variance is fixed to a multiple of identity: $\forall i,
\Sigma_i=\sigma^2 I_v$ where $\sigma$ is an hyper-parameter to choose.

@@

The complexity of Deterministic SRM is in $\bigO(\mathrm{n_{iter}} mpvn)$ where
$n$ is the number of samples and $\mathrm{n_{iter}}$ the number of iterations.
%
We monitor the convergence by looking at the $\ell_{\infty}$ norm of the
gradient.
%
Note that we can monitor the gradient without any increase in complexity.
%
Indeed, after the updates with respect to each mixing matrix have been
carried out, the gradient with respect to $(W_i)_{i=1}^m$ is zero and therefore
to compute the $\ell_{\infty}$ norm of the gradient we only need the gradient
with respect to $S$: $m S - \sum_i A_i^{\top}\xb_i$ where the right hand side is
used in the updates of $S$.
%
The algorithm is stopped when the
gradient falls below a chosen tolerance.
%

@@

The model is optimized via the expectation maximization algorithm.
%
Denoting $\VV[\sbb | \xb] = (\sum_i \frac1{\sigma_i^2} I +
\Sigma_s^{-1})^{-1}$ and $\EE[\sbb | \xb] = \VV[\sbb | \xb] \sum_i \frac1{\sigma_i^2}
A_i^{\top}\xb_i$, we have
\begin{align}
  p(\xb, \sbb) &= \prod_i \frac{\exp \left(-\frac{\|\xb_i - A_i \sbb \|^2}{2 \sigma_i^2} \right) }{(2 \pi \sigma_i^{2v})^{\frac12}} \frac{\exp(-\frac12 \dotp{ \sbb }{ \Sigma_s^{-1} \sbb } )}{(2 \pi | \Sigma_s|)^{\frac12}} \\
               &= c_1 \exp (-\frac12 \left( \sum_i \frac1{\sigma_i^2}\|\xb_i\|^2 - 2  \dotp{ \sum_i \frac1{\sigma_i^2} A_i^{\top}\xb_i}{ \sbb } \right. \\& \left.+ \sum_i \frac1{\sigma_i^2} \| \sbb \|^2 + \dotp{ \sbb}{ \Sigma_s^{-1} \sbb }  \right)) \\
               &= c_2(\xb) \exp(-\frac12 \left( \dotp{  \sbb - \EE[\sbb }{ \xb], \VV[\sbb | \xb]^{-1} ( \sbb - \EE[\sbb | \xb])  } \right)) \label{probsrmcompleted}
\end{align}
where
\begin{align}
  c_1 = \frac1{(2 \pi \sigma_i^{2v})^{\frac12}}\frac1{(2 \pi |
  \Sigma_s|)^{\frac12}}
\end{align}
and
\begin{align}
  c_2(\xb) = c_1 \exp \left( -\frac12( \sum_i
  \frac1{\sigma_i^2}\|\xb_i\|^2 - \dotp{  \EE[\sbb }{ \xb], \VV[\sbb | \xb]^{-1}
  \EE[\sbb | \xb] }) \right)
\end{align}
are independent from $\sbb$.
%
Therefore,
\begin{align}
  \sbb| \xb \sim \Ncal(\EE[\sbb | \xb], \VV[\sbb, \xb])
\end{align}

@@

After preprocessing, the reduced representation $\zb_i$ is used instead of the
original data $\xb_i$ yielding a time complexity of $\bigO(
\mathrm{T_{preprocessing}} + \mathrm{n_{iter}} mpnr)$.
%
Let us highlight that an experiment is often run multiple times,
typically when the analysis requires cross-validation procedures.
%
In these cases,
the pre-processing is performed only once and the apparent complexity becomes
$\bigO(\mathrm{n_{iter}} mpnr)$ which is faster than SRM by
a factor of $\frac{v}{r}$.
%
The number of components in large spatial decompositions is about $r=1000$ and in full brain data,
the number of voxels is about $300~000$ so that $\frac{v}{r}$ is typically about
$300$.
%
It remains to be shown how to draw a correspondence between FastSRM and SRM which is
addressed in the following section.
%

@@

A complexity analysis shows that finding an optimal spatial decomposition becomes the limiting
step of the pipeline.
%
Even with fast implementations, subject specific PCA
is costly.
%
However FastSRM only works on $\zb_i$ so we do not need to know the value of
$U_{\xb_i}$.
%
In practice, we observe data $X_i \in \RR^{v \times n}$ and we want to get $Z_i
\in \RR^{n \times n}$ such that $X_i = U_{\xb_i} Z_i$.
%
This can be done by performing an
SVD of $X_i^{\top} X_i$ yielding $X_i^{\top}X_i= V_i D_i V_i^{\top}$ and setting
$Z_i = D_i^{\frac12} V_i^{\top}$.
%
Although computing the product $X_i^{\top} X_i$ has time complexity
$\bigO(vt^2)$ (the same as a PCA) the constant in the $\bigO$ is exactly one so
this operation costs a lot less than a PCA
on full data.
%
When estimates of the mixing matrices are needed, they can be obtained by
applying equation~\eqref{eq:fastdetsrm:Aiupdate} in the deterministic SRM case and
equation~\eqref{eq:fastprobsrm:Aiupdate} in the probabilistic SRM case which only costs
$\bigO(mvp^2)$.
%
In practice the cost of the matrix products $X_i^{\top} X_i$ is often still the
limiting step of the pipeline (this depends on the number of iterations) but as
we show next, it is much more efficient than performing SRM on
the full data.
%
Note that if memory allows it, these matrix products can be
computed in parallel.
%

\section{Related work}
The implementation of SRM proposed in~\cite{chen2015reduced} is quadratic in the
number of features which prevent its application to full brain data.
In~\cite{anderson2016enabling}, this issue is addressed by using the
inversion lemma to remove the quadratic dependence. Their implementation is now the most widely used one.
In our work, we propose to reduce further the computation time and memory usage
compared to the implementation in~\cite{anderson2016enabling} by the use of
optimal spatial decompositions.

@@

We see SRM as a well principled formulation of the hyperalignment
method~\cite{haxby2011common}.
%
Many methods exist to perform hyperalignment: deep
hyperalignment~\cite{yousefnezhad2017deep} , robust
SRM~\cite{turek2018capturing} , multi-view canonical correlation
analysis~\cite{li2009joint}, ShICA~\cite{richard2021model},
MultiViewICA~\cite{richard2020modeling} optimal
transport~\cite{bazeille2019local} and many more.
%
In this work, we do not discuss the merit of FastSRM with respect to
other methods. We only see that SRM is widely used and aim to provide
a faster, more memory efficient and identifiable implementation.  In
our view FastSRM should be seen as a plugin replacement for SRM.

\section{Experiments}
We make several experiments on both synthetic and real data.
%
We used nilearn~\cite{abraham2014machine} (version 0.8.1) for fMRI data preprocessing,
Brainiak~\cite{kumar2020brainiak2} (version 0.9) for the non-identifiable version of SRM that
implements the work of~\cite{chen2015reduced} and~\cite{anderson2016enabling},
Numpy~\cite{harris2020array} (version 1.19.0) for
array processing, Scipy~\cite{2020SciPy-NMeth} (version 1.5.3) its implementation of the
Hungarian algorithm, Matplotlib~\cite{hunter2007matplotlib} (version 3.1.2) for plotting and
Sklearn~\cite{pedregosa2011scikit} (version 0.23.2) for machine learning pipelines.

@@

\subsection{Identifiability increases stability}
Assuming that the data follow the model, identifiability ensures that the
problem is well posed since the solutions are well-characterized. Note however that SRM
remains non-convex so there is no guarantee that the optimal solution is found.
%
% \bt{I find this a bit ambiguous: you give te impression that you gurantee that the optimal solution is used. Rephrase.}
%
In real fMRI data, the model cannot be expected to hold perfectly, but we can
hope for greater stability in the parameters recovered than if an unidentifiable
model is used.

@@

While brainiak and nilearn are useful to reproduce some of our experiments, they
are not strong dependencies and FastSRM only needs numpy ($\geq 1.12$), scipy
($\geq 0.18.0$), matplotlib ($\geq 2.0.0$), scikit-learn ($\geq 0.23$), pytest
($\geq 6.2.5$) , joblib ($\geq 1.1.0$).

@@

\section{Conclusion}
As studies using naturalistic stimuli will tend to become more common and large
within and across subjects, we need scalable models in terms of computation time
and memory usage.
%
This is particularly related to the development of deep phenotyping
models such as the Courtois project on Neural Modeling (see \url{https://docs.cneuromod.ca}).
%
This is what FastSRM provides.

@@

%
FastSRM inherits from all the weaknesses of SRM including the fact that
mixing matrices are assumed to be orthogonal, which is rather unrealistic. It
remains to be seen whether dimension reduction and learning of the shared
components could be done jointly and efficiently without assuming orthogonal
mixing matrices.
%
Our approach based on optimal spatial decomposition shows that SRM-like methods do not need the
full data to provide accurate estimates of shared components. We believe such
insights may guide the design of future methods.
%

@@

  Then we consider the expected negative log-likelihood of probabilistic srm:
  \begin{align}
    \loss &= \sum_i \frac12 v\log(\sigma_i^2) + \frac12 \log(|\Sigma_s|) + \EE[\int_{\sbb} \sum_i \frac1{2 \sigma_i^{2}} \|\xb_i - A_i \sbb \|^2 \nonumber \\& \enspace \enspace \enspace \enspace+ \frac12 \dotp{ \sbb }{ \Sigma_s^{-1} \sbb }  d\sbb] \\
          &= \sum_i \frac12 v \log(\sigma_i^2) + \frac12 \log(|\Sigma_s|) + \EE[\int_{\sbb} \sum_i \frac1{2 \sigma_i^{2}}\|\zb_i - \tilde{A}_i \sbb \|^2 \nonumber \\& \enspace \enspace \enspace \enspace+ \frac12 \dotp{ \sbb }{ \Sigma_s^{-1} \sbb }  d\sbb]
  \end{align}
  where we use equality~\eqref{eq:equality:xy}.
  %

  Optimizing the log-likelihood via expectation maximization yields the exact
  same updates as probabilistic SRM on data $\zb_i$
  except that updates
  \begin{align}
    \sigma_i^2 \leftarrow \frac1{t} (\EE[\| \zb_i - \tilde{A}_i \EE[\sbb|\zb]\|^2] + \| \diag(\VV[\sbb | \zb]) \|^2)
  \end{align}
  are replaced by updates
  \begin{align}
    \sigma_i^2 \leftarrow \frac1{v} (\EE[\| \zb_i - \tilde{A}_i \EE[\sbb|\zb]\|^2] + \| \diag(\VV[\sbb | \zb]) \|^2)
  \end{align}
  so that $\tilde{A}_i = A'_i$.
  %

  Therefore, the updates in both algorithms are linked  by $A_i = U_{\xb_i}A_i'$,
  $\sigma_i' =
  \sigma_i$ and $\Sigma_s'  = \Sigma_s$.
  %

  This concludes the proof.
  %

\end{proof}
\end{document}

>>>>

In this work, we implement an identifiable version of SRM and show on real data
that it improves the stability of the recovered shared response.
%
We then introduce FastSRM, that relies on a dimension reduction step that we
prove to yield the same solution as the original algorithm.
% be optimal in the sense that working with reduced data does not induce any change in the
% algorithm trajectory.
%
We show experimentally using synthetic and real fMRI data
that FastSRM is considerably faster and more memory efficient
than current implementations.

@@

\section{Introduction}
%\bt{We need to start with a more application-oriented intro}
%
Complex stimulations such as movie-watching, story or music listening are more
and more popular in the neuro-scientific community. Indeed such naturalistic
paradigms are unconstrained from behavioral manipulations and thus, more
ecological with respect to every-day cognitive conditions.
%
Yet, in contrast to standard designs, these so-called
naturalistic stimuli do not come with a design matrix that quantifies the
features of interest in the stimuli.
%
Without a design matrix, models such as the general linear
model~\cite{poline2012general} cannot be used.
%
Although some works have used manual
annotations~\cite{huth2012continuous}, see also
\url{https://neuroscout.org} or deep neural
networks~\cite{gucclu2017increasingly}~\cite{richard2018optimizing} to
create an estimate of the design matrix of naturalistic stimuli it is
a hard task.

Another approach is to learn jointly the design matrix and the spatial maps in
an unsupervised way.
%
In the case where the dataset contains only one subject,
independent component analysis~\cite{jutten1991blind} (ICA) is the method of choice.
%
ICA assumes that components are independent, a defendible assumption which
ensures identifiability up to permutation and scaling of the model and can be fitted efficiently.
%

But when the dataset contains multiple subjects, a natural assumption is to
assume that the design matrix is shared across subjects.
%
Many different methods
can produce shared components from different subjects.
%
Some assume independent
components~\cite{richard2021model}~\cite{richard2020modeling}~\cite{varoquaux2009canica}~\cite{calhoun2001method},
while others are only based on second order information such as
Multiset Canonical Correlation Analysis~\cite{via2011joint} or the
Shared Response Model (SRM)~\cite{chen2015reduced}.
%

This work focuses on the SRM. Because of its simplicity and its
built-in dimension reduction, this model is widely
used~\cite{baldassano2017discovering}~\cite{cohen2017computational}~\cite{baldassano2018representation}~\cite{jolly2020custom}~\cite{lee2021anticipation},
in particular, as a preprocessing step for source
identification~\cite{richard2021model}~\cite{richard2020modeling},
classification~\cite{turek2018capturing}~\cite{chen2017shared}\cite{zhang2016searchlight}
or as a basis for transfer learning~\cite{zhang2018transfer}.

@@

In this work, we introduce FastSRM, an identifiable model that uses an optimal
spatial decomposition to speed up the computations with provably no loss of
performance.
%
FastSRM gains identifiability by imposing that the covariance of the shared
components is diagonal (see ~\cite{richard2020modeling} Appendix D for a proof).
%
The gains in speed and memory usage are based on the observation that a low-dimensional, yet distance-preserving
representation of the images yields the same result as the full data.
%
Such a representation can be interpreted as an initial spatial  decomposition of the data.
% that reduces the data in a such a way that the obtained
% shared components (and other parameters) are the same as if they were computed
% on full data.
%
We show that FastSRM brings several practical benefits: on real data, it produces more stable
estimates of the shared components than its non-identifiable counterpart, and 
 it is much faster and more memory
efficient than other implementations that do not make use of an optimal spatial decomposition.
% on synthetic as well as on real data,

@@

\subsection{Deterministic shared response model}
\label{sec:deterministicsrm}
Let us consider $n$ observations (scans) of $\xb_i$ and $\sbb$ that we stack into
matrices $X_i \in \RR^{v, n}$ and $S \in \RR^{p, n}$.
%
The deterministic shared response model considers both the mixing matrices $A_i$ and
the $n$ observations of the shared response $S$ as parameters to be
estimated.
%
The noise variance is fixed to a multiple of identity: $\forall i,
\Sigma_i=\sigma^2 I_v$ where $\sigma$ is an hyper-parameter to choose.

@@

The complexity of Deterministic SRM is in $\bigO(\mathrm{n_{iter}} mpvn)$, where
 $\mathrm{n_{iter}}$ is the number of iterations.
%
We monitor the convergence by computing the $\ell_{\infty}$ norm of the
gradient.
%
Note that we can monitor the gradient without any increase in complexity.
%
Indeed, after the updates with respect to each mixing matrix have been
carried out, the gradient with respect to $(W_i)_{i=1}^m$ is zero and therefore
to compute the $\ell_{\infty}$ norm of the gradient we only need the gradient
with respect to $S$: $m S - \sum_i A_i^{\top}\xb_i$ where the right hand side is
used in the updates of $S$.
%
The algorithm is stopped when the
gradient falls below a chosen tolerance.
%

@@

The model is optimized via the expectation maximization algorithm.
%
Denoting $\VV[\sbb | \xb] = (\sum_i \frac1{\sigma_i^2} I +
\Sigma_s^{-1})^{-1}$ and $\EE[\sbb | \xb] = \VV[\sbb | \xb] \sum_i \frac1{\sigma_i^2}
A_i^{\top}\xb_i$, we have
\begin{align}
  p(\xb, \sbb) &= \prod_i \frac{\exp \left(-\frac{\|\xb_i - A_i \sbb \|^2}{2 \sigma_i^2} \right) }{(2 \pi \sigma_i^{2v})^{\frac12}} \frac{\exp(-\frac12 \dotp{ \sbb }{ \Sigma_s^{-1} \sbb } )}{(2 \pi | \Sigma_s|)^{\frac12}} \\
               &= c_1 \exp (-\frac12 \left( \sum_i \frac1{\sigma_i^2}\|\xb_i\|^2 - 2  \dotp{ \sum_i \frac1{\sigma_i^2} A_i^{\top}\xb_i}{ \sbb } \right. \\& \left.+ \sum_i \frac1{\sigma_i^2} \| \sbb \|^2 + \dotp{ \sbb}{ \Sigma_s^{-1} \sbb }  \right)) \\
               &= c_2(\xb) \exp(-\frac12 \left( \dotp{  \sbb - \EE[\sbb }{ \xb], \VV[\sbb | \xb]^{-1} ( \sbb - \EE[\sbb | \xb])  } \right)) \label{probsrmcompleted}
\end{align}
where
\begin{align}
  c_1 = \frac1{(2 \pi \sigma_i^{2v})^{\frac12}}\frac1{(2 \pi |
  \Sigma_s|)^{\frac12}}
\end{align}
and
\begin{align}
  c_2(\xb) = c_1 \exp \left( -\frac12 \left( \sum_i \frac1{\sigma_i^2}\|\xb_i\|^2 - \dotp{  \EE[\sbb }{ \xb], \VV[\sbb | \xb]^{-1} \EE[\sbb | \xb] } \right) \right)
\end{align}
are independent from $\sbb$.
%
Therefore,
\begin{align}
  \sbb| \xb \sim \Ncal(\EE[\sbb | \xb], \VV[\sbb, \xb])
\end{align}

@@

After preprocessing, the reduced representation $\zb_i$ is used instead of the
original data $\xb_i$ yielding a time complexity of $\bigO(
\mathrm{T_{preprocessing}} + \mathrm{n_{iter}} mpnr)$.
%
Let us highlight that an experiment is often run multiple times,
typically when the analysis requires cross-validation procedures.
%
In these cases,
the pre-processing is performed only once and the apparent complexity becomes
$\bigO(\mathrm{n_{iter}} mpnr)$ which is faster than SRM by
a factor of $\frac{v}{r}$.
%
The number of components in large spatial decompositions is about $r=1000$ and in full brain data,
the number of voxels is about $300~000$ so that $\frac{v}{r}$ is typically about
$300$.
%
It remains to be shown how to draw a correspondence between FastSRM and SRM, which is
addressed in the following section.
%

@@

A complexity analysis shows that finding an optimal spatial decomposition becomes the limiting
step of the pipeline.
%
Even with fast implementations, subject specific PCA
is costly.
%
However FastSRM only works on $\zb_i$ so we do not need to know the value of
$U_{\xb_i}$.
%
In practice, we observe data $X_i \in \RR^{v \times n}$ and we want to get $Z_i
\in \RR^{n \times n}$ such that $X_i = U_{\xb_i} Z_i$.
%
This can be done by performing an
SVD of $X_i^{\top} X_i$ yielding $X_i^{\top}X_i= V_i D_i V_i^{\top}$ and setting
$Z_i = D_i^{\frac12} V_i^{\top}$.
%
Although computing the product $X_i^{\top} X_i$ has time complexity
$\bigO(vt^2)$ (the same as a PCA) the constant in the $\bigO$ is exactly one so
this operation costs a lot less than a PCA
on full data.
%
When estimates of the mixing matrices are needed, they can be obtained by
applying equation~\eqref{eq:fastdetsrm:Aiupdate} in the deterministic SRM case and
equation~\eqref{eq:fastprobsrm:Aiupdate} in the probabilistic SRM case which only costs
$\bigO(mvp^2)$.
%
In practice the cost of the matrix products $X_i^{\top} X_i$ is often the
limiting step of the pipeline (this depends on the number of iterations) but as
we show next, it is much more efficient than performing SRM on
the full data.
%
Note that if large  memory resources are available, these matrix products can be
computed in parallel.
%

\section{Related work}
The implementation of SRM proposed in~\cite{chen2015reduced} is quadratic in the
number of features, which prevents its application to full brain data.
In~\cite{anderson2016enabling}, this issue is addressed by using the
inversion lemma to remove the quadratic dependence. Their implementation is now the most widely used one.
In our work, we propose to reduce further the computation time and memory usage
compared to the implementation in~\cite{anderson2016enabling} by the use of
optimal spatial decompositions.

@@

We consider SRM as a well principled formulation of the hyperalignment
method~\cite{haxby2011common}.
%
Many methods exist to perform hyperalignment: deep
hyperalignment~\cite{yousefnezhad2017deep} , robust
SRM~\cite{turek2018capturing} , multi-view canonical correlation
analysis~\cite{li2009joint}, ShICA~\cite{richard2021model},
MultiViewICA~\cite{richard2020modeling} optimal
transport~\cite{bazeille2019local} and many more.
%
In this work, we do not discuss the merit of FastSRM with respect to
other methods. We simply observe that SRM is widely used and aim to provide
a faster, more memory efficient and identifiable implementation.  In
our view FastSRM, should be used as a plugin replacement for SRM.

\section{Experiments}
We make several experiments on both synthetic and real data.
%
We used Nilearn~\cite{abraham2014machine} (version 0.8.1) for fMRI data preprocessing,
Brainiak~\cite{kumar2020brainiak2} (version 0.9) for the non-identifiable version of SRM that
implements the work of~\cite{chen2015reduced} and~\cite{anderson2016enabling},
Numpy~\cite{harris2020array} (version 1.19.0) for
array processing, Scipy~\cite{2020SciPy-NMeth} (version 1.5.3) its implementation of the
Hungarian algorithm, Matplotlib~\cite{hunter2007matplotlib} (version 3.1.2) for plotting and
Sklearn~\cite{pedregosa2011scikit} (version 0.23.2) for machine learning pipelines.

@@

\subsection{Identifiability increases stability}
Assuming that the data follow the model, identifiability ensures that the
problem is well posed since the solutions are well-characterized. Note however that SRM
remains non-convex so there is no guarantee that the optimal solution is found.
%
% \bt{I find this a bit ambiguous: you give te impression that you gurantee that the optimal solution is used. Rephrase.}
%
In real fMRI data, the model cannot be expected to hold perfectly, but we can
hope for greater stability in the parameters recovered than if an non-identifiable
model is used.

@@

While brainiak and nilearn are useful to reproduce some of our experiments, they
are not strong dependencies and FastSRM only needs Numpy ($\geq 1.12$), Scipy
($\geq 0.18.0$), Matplotlib ($\geq 2.0.0$), Scikit-learn ($\geq 0.23$), Pytest
($\geq 6.2.5$) , and Joblib ($\geq 1.1.0$).

@@

\section{Conclusion}
As studies using naturalistic stimuli will tend to become more common and large
within and across subjects, we need scalable models in terms of computation time
and memory usage.
%
This is particularly related to the development of deep phenotyping
models such as the Courtois project on Neural Modeling (see \url{https://docs.cneuromod.ca}).
%
This is precisely what FastSRM provides.

@@

%
FastSRM inherits from all the weaknesses of SRM including the fact that
mixing matrices are assumed to be orthogonal, which is debattable in practice. It
remains to be seen whether dimension reduction and learning of the shared
components could be done jointly and efficiently without assuming orthogonal
mixing matrices.
%
Our approach based on optimal spatial decomposition shows that SRM-like methods do not need the
full data to provide accurate estimates of shared components. We believe such
insights may guide the design of future methods.
%

@@

  Then we consider the expected negative log-likelihood of probabilistic srm:
  \begin{align}
    \loss &= \sum_i \frac12 v\log(\sigma_i^2) + \frac12 \log(|\Sigma_s|) + \EE[\int_{\sbb} \sum_i \frac1{2 \sigma_i^{2}} \|\xb_i - A_i \sbb \|^2 \nonumber \\& \enspace \enspace \enspace \enspace+ \frac12 \dotp{ \sbb }{ \Sigma_s^{-1} \sbb }  d\sbb] \\
          &= \sum_i \frac12 v \log(\sigma_i^2) + \frac12 \log(|\Sigma_s|) + \EE[\int_{\sbb} \sum_i \frac1{2 \sigma_i^{2}}\|\zb_i - \tilde{A}_i \sbb \|^2 \nonumber \\& \enspace \enspace \enspace \enspace+ \frac12 \dotp{ \sbb }{ \Sigma_s^{-1} \sbb }  d\sbb]
  \end{align}
  where we use equality~\eqref{eq:equality:xy}.
  %
  %
  Optimizing the log-likelihood via expectation maximization yields the exact
  same updates as probabilistic SRM on data $\zb_i$
  except that updates
  \begin{align}
    \sigma_i^2 \leftarrow \frac1{t} (\EE[\| \zb_i - \tilde{A}_i \EE[\sbb|\zb]\|^2] + \| \diag(\VV[\sbb | \zb]) \|^2)
  \end{align}
  are replaced by updates
  \begin{align}
    \sigma_i^2 \leftarrow \frac1{v} (\EE[\| \zb_i - \tilde{A}_i \EE[\sbb|\zb]\|^2] + \| \diag(\VV[\sbb | \zb]) \|^2)
  \end{align}
  so that $\tilde{A}_i = A'_i$.
  %
  Therefore, the updates in both algorithms are linked  by $A_i = U_{\xb_i}A_i'$,
  $\sigma_i' =
  \sigma_i$ and $\Sigma_s'  = \Sigma_s$.
  %
  This concludes the proof.
  %
\end{proof}
\end{document}
